---
title: "DeNovoWEST weight creation"
author: "Joanna Kaplanis & Kaitlin E. Samocha"
date: "05/06/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction  

This is a guide for how to determine the positive predictive value (PPV) weights that are used in [DeNovoWEST](https://www.biorxiv.org/content/10.1101/797787v3). To load files, you will need to provide the filepaths, specifically the `rates_directory` and the full path to the _de novo_ mutation file (`dnm_file_wweights`) that has been annotated with CADD scores, sHet high vs low, and whether the variant falls into a region of missense constraint. As written, it assumes that the provided files, with no changes in name, are being loaded.    

This code was written by Joanna Kaplanis (@queenjobo) and Kaitlin Samocha (@ksamocha).  

```{r set up}
library(data.table)
library(dplyr)
library(ggplot2)
library(wesanderson)
mypal <- wes_palette("Royal1")

rates_directory = 'path_here'
dnm_file_wweights = 'path_to_file_here/DDD_RUMC_GDX_denovos_cadd_shet_wweights_2020_01_17.txt.gz'
```

```{r load files}
# load dnms and add size of the variant
all_wweights <- fread(dnm_file_wweights, stringsAsFactors = F, sep = '\t', header = T)
all_wweights$size <- nchar(all_wweights$alt) - nchar(all_wweights$ref)

# load rates -- takes a while to load!
rates_1 <- fread(paste(rates_directory, 'all_rates_0_4000_cadd_maf_shet_2020_01_25.txt.gz', sep=''), sep = '\t', header = T, stringsAsFactors = F)
rates_2 <- fread(paste(rates_directory, 'all_rates_4000_8000_cadd_maf_shet_2020_01_25.txt.gz', sep=''), sep = '\t', header = T, stringsAsFactors = F)
rates_3 <- fread(paste(rates_directory, 'all_rates_8000_12000_cadd_maf_shet_2020_01_25.txt.gz', sep=''), sep = '\t', header = T, stringsAsFactors = F)
rates_4 <- fread(paste(rates_directory, 'all_rates_12000_16000_cadd_maf_shet_2020_01_25.txt.gz', sep=''), sep = '\t', header = T, stringsAsFactors = F)
rates_5 <- fread(paste(rates_directory, 'all_rates_16000_20000_cadd_maf_shet_2020_01_25.txt.gz', sep=''), sep = '\t', header = T, stringsAsFactors = F)
```

## Make weights  

This is set up for analysis of the 31,058 individuals included as part of Kaplanis, Samocha, Wiel, Zhang et al. If running on another dataset, the number of males (`male_n`) and females (`female_n`) will need to be adjusted.  

```{r determine chrX factor}
# determine X factor to correct chrX -- taken from Jeremy McRae's code on github (https://github.com/jeremymcrae/mupit/blob/master/mupit/mutation_rates.py)
male_n = 17422
female_n = 13636
autosomal = 2 * (male_n + female_n)
female_transmissions = male_n + female_n
male_transmissions = female_n
    
# get scaling factors using the alpha from the most recent SFHS (Scottish
# Family Health Study) phased de novo data.
alpha = 3.4
male_factor = 2 / (1 + (1 / alpha))
female_factor = 2 / (1 + alpha)
    
# correct the non-PAR chrX genes for fewer transmissions and lower rate
# (dependent on alpha)
x_factor = ((male_transmissions * male_factor) + (female_transmissions * female_factor)) / autosomal
```

### Synonymous weights  

For synonymous variants, we split by whether the variant is in a high sHet gene (sHet >= 0.15) or a low sHet gene.  

```{r functions for 2 bins}
get_expected_by_shet2bins_bycq <- function(rate_group, givencq) {
  return_rates <- rate_group %>%
    filter(cq==givencq) %>%
    mutate(sex_chr = ifelse(chrom=='X', 'sex', 'auto')) %>%
    mutate(shetbin = ifelse(shethigh=='TRUE', 'high', 'low')) %>%
    group_by(shetbin, sex_chr) %>%
    summarize(sum_prob = sum(prob)) %>%
    mutate(expected = ifelse(sex_chr=='auto', sum_prob*autosomal, sum_prob*autosomal*x_factor)) %>%
    group_by(shetbin) %>%
    summarize(total_expected = sum(expected))

  return(return_rates)
}
```

```{r synonymous}
## need to determine enrichments by shet categories
obs_syn_shet2bins <- all_wweights %>%
  filter(cq=='synonymous_variant' & size==0) %>%
  mutate(shetbin = ifelse(shethigh=='TRUE', 'high', 'low')) %>%
  group_by(shetbin) %>%
  tally()

shet2bins_syn_exp_rates1 <- get_expected_by_shet2bins_bycq(rates_1, 'synonymous')
shet2bins_syn_exp_rates2 <- get_expected_by_shet2bins_bycq(rates_2, 'synonymous')
shet2bins_syn_exp_rates3 <- get_expected_by_shet2bins_bycq(rates_3, 'synonymous')
shet2bins_syn_exp_rates4 <- get_expected_by_shet2bins_bycq(rates_4, 'synonymous')
shet2bins_syn_exp_rates5 <- get_expected_by_shet2bins_bycq(rates_5, 'synonymous')

combined_shet2bins_syn_exp_rates <- bind_rows(shet2bins_syn_exp_rates1,
                                              shet2bins_syn_exp_rates2,
                                              shet2bins_syn_exp_rates3,
                                              shet2bins_syn_exp_rates4,
                                              shet2bins_syn_exp_rates5) %>%
  replace(., is.na(.), 0) %>%
  group_by(shetbin) %>%
  summarize_all(sum)

obs_exp_shet2bins_syn <- obs_syn_shet2bins %>%
  left_join(combined_shet2bins_syn_exp_rates, by=c('shetbin')) %>%
  mutate(obs_exp = n/total_expected)

# add confidence intervals
obs_exp_shet2bins_syn$ci.lower <- sapply(obs_exp_shet2bins_syn$n, function(x) poisson.test(x, conf.level = 0.95)$conf.int[1])/obs_exp_shet2bins_syn$total_expected
obs_exp_shet2bins_syn$ci.upper <- sapply(obs_exp_shet2bins_syn$n, function(x) poisson.test(x, conf.level = 0.95)$conf.int[2])/obs_exp_shet2bins_syn$total_expected

synon <- data.frame(cq = "synonymous",
                    score = NA,
                    shetbin = c('high', 'low'),
                    con = NA,
                    conlower = NA,
                    conupper = NA,
                    uncon = obs_exp_shet2bins_syn %>% pull(obs_exp),
                    unconlower = obs_exp_shet2bins_syn %>% pull(ci.lower),
                    unconupper = obs_exp_shet2bins_syn %>% pull(ci.upper)
                  )
```

### Missense weights  

For this work, we split missense variants by sHet (high vs low), missense constraint (in a region vs not in a region of missense constraint), and by six bins for CADD scores. We found similar results for a different number of CADD bins. This is the section that will need to be modified heavily for different datasets (e.g. fewer bins if the datasets are smaller).  

```{r function for 6 bins of cadd and missense}
get_expected_by_shet2bins_cadd6bins <- function(rate_group) {
  return_rates <- rate_group %>%
    filter(cq=='missense') %>%
    mutate(sex_chr = ifelse(chrom=='X', 'sex', 'auto')) %>%
    mutate(shetbin = ifelse(shethigh=='TRUE', 'high', 'low')) %>%
  mutate(cadd_bin = case_when(score < 6 ~ 1,
                              score < 12 & score >= 6 ~ 2,
                              score < 18 & score >= 12 ~ 3,
                              score < 24 & score >= 18 ~ 4,
                              score < 30 & score >= 24 ~ 5,
                              score >= 30 ~ 6)) %>%
    filter(!is.na(cadd_bin)) %>%
    group_by(cadd_bin, shetbin, constrained, sex_chr) %>%
    summarize(sum_prob = sum(prob)) %>%
    mutate(expected = ifelse(sex_chr=='auto', sum_prob*autosomal, sum_prob*autosomal*x_factor)) %>%
    group_by(cadd_bin, shetbin, constrained) %>%
    summarize(total_expected = sum(expected))

  return(return_rates)
}
```

```{r missense}
shet2bins_cadd6bins_missense_obs <- all_wweights %>%
  filter(cq=='missense_variant') %>%
  mutate(shetbin = ifelse(shethigh=='TRUE', 'high', 'low')) %>%
  mutate(cadd_bin = case_when(score < 6 ~ 1,
                              score < 12 & score >= 6 ~ 2,
                              score < 18 & score >= 12 ~ 3,
                              score < 24 & score >= 18 ~ 4,
                              score < 30 & score >= 24 ~ 5,
                              score >= 30 ~ 6)) %>%
  filter(!is.na(cadd_bin)) %>%
  group_by(cadd_bin, shetbin, constrained) %>%
  tally()

shet2bins_cadd6bins_missense_exp_rates1 <- get_expected_by_shet2bins_cadd6bins(rates_1)
shet2bins_cadd6bins_missense_exp_rates2 <- get_expected_by_shet2bins_cadd6bins(rates_2)
shet2bins_cadd6bins_missense_exp_rates3 <- get_expected_by_shet2bins_cadd6bins(rates_3)
shet2bins_cadd6bins_missense_exp_rates4 <- get_expected_by_shet2bins_cadd6bins(rates_4)
shet2bins_cadd6bins_missense_exp_rates5 <- get_expected_by_shet2bins_cadd6bins(rates_5)

combined_shet2bins_cadd6bins_missense_exp_rates <- bind_rows(shet2bins_cadd6bins_missense_exp_rates1,
                                                             shet2bins_cadd6bins_missense_exp_rates2,
                                                             shet2bins_cadd6bins_missense_exp_rates3,
                                                             shet2bins_cadd6bins_missense_exp_rates4,
                                                             shet2bins_cadd6bins_missense_exp_rates5) %>%
  replace(., is.na(.), 0) %>%
  group_by(cadd_bin, shetbin, constrained) %>%
  summarize_all(sum)

obs_exp_shet2bins_cadd6bins_missense <- shet2bins_cadd6bins_missense_obs %>%
  left_join(combined_shet2bins_cadd6bins_missense_exp_rates, by=c('cadd_bin', 'shetbin', 'constrained')) %>%
  mutate(obs_exp = n/total_expected)

# add confidence intervals
obs_exp_shet2bins_cadd6bins_missense$ci.lower <- sapply(obs_exp_shet2bins_cadd6bins_missense$n, function(x) poisson.test(x, conf.level = 0.95)$conf.int[1])/obs_exp_shet2bins_cadd6bins_missense$total_expected
obs_exp_shet2bins_cadd6bins_missense$ci.upper <- sapply(obs_exp_shet2bins_cadd6bins_missense$n, function(x) poisson.test(x, conf.level = 0.95)$conf.int[2])/obs_exp_shet2bins_cadd6bins_missense$total_expected

ggplot(obs_exp_shet2bins_cadd6bins_missense, aes(x = cadd_bin, y = obs_exp, color=shetbin, shape=constrained, lty=constrained)) +
  geom_point() +
  geom_line() +
  geom_pointrange(aes(ymin=ci.lower, ymax=ci.upper)) +
  theme_classic() +
  geom_hline(yintercept = 1, color='gray', lty=2) +
  labs(x = "CADD", y = 'Enrichment (obs/exp missense)') +
  scale_color_manual(values = c(mypal[2], mypal[1])) +
  scale_x_continuous(name = 'CADD', breaks=seq(1,6,1), labels=c('0-6', '6-12', '12-18', '18-24', '24-30', '30+'))
```

```{r run loess for missense variants}
# missense -- 4 categories
forloess_missense_shet2bins_cadd6bins_high_con <- obs_exp_shet2bins_cadd6bins_missense %>%
  filter(shetbin=='high' & constrained=='TRUE') %>%
  mutate(midpoint = case_when(cadd_bin==1 ~ 3+0,
                              cadd_bin==2 ~ 3+6,
                              cadd_bin==3 ~ 3+12,
                              cadd_bin==4 ~ 3+18,
                              cadd_bin==5 ~ 3+24,
                              cadd_bin==6 ~ 3+30))
forloess_missense_shet2bins_cadd6bins_high_uncon <- obs_exp_shet2bins_cadd6bins_missense %>%
  filter(shetbin=='high' & constrained=='FALSE') %>%
  mutate(midpoint = case_when(cadd_bin==1 ~ 3+0,
                              cadd_bin==2 ~ 3+6,
                              cadd_bin==3 ~ 3+12,
                              cadd_bin==4 ~ 3+18,
                              cadd_bin==5 ~ 3+24,
                              cadd_bin==6 ~ 3+30))
forloess_missense_shet2bins_cadd6bins_low_con <- obs_exp_shet2bins_cadd6bins_missense %>%
  filter(shetbin=='low' & constrained=='TRUE') %>%
  mutate(midpoint = case_when(cadd_bin==1 ~ 3+0,
                              cadd_bin==2 ~ 3+6,
                              cadd_bin==3 ~ 3+12,
                              cadd_bin==4 ~ 3+18,
                              cadd_bin==5 ~ 3+24,
                              cadd_bin==6 ~ 3+30))
forloess_missense_shet2bins_cadd6bins_low_uncon <- obs_exp_shet2bins_cadd6bins_missense %>%
  filter(shetbin=='low' & constrained=='FALSE') %>%
  mutate(midpoint = case_when(cadd_bin==1 ~ 3+0,
                              cadd_bin==2 ~ 3+6,
                              cadd_bin==3 ~ 3+12,
                              cadd_bin==4 ~ 3+18,
                              cadd_bin==5 ~ 3+24,
                              cadd_bin==6 ~ 3+30))

results.loess_high_missense_con <- loess(forloess_missense_shet2bins_cadd6bins_high_con$obs_exp ~ forloess_missense_shet2bins_cadd6bins_high_con$midpoint, weights = forloess_missense_shet2bins_cadd6bins_high_con$n, span=0.99)
results.loess_low_missense_con <- loess(forloess_missense_shet2bins_cadd6bins_low_con$obs_exp ~ forloess_missense_shet2bins_cadd6bins_low_con$midpoint, weights = forloess_missense_shet2bins_cadd6bins_low_con$n, span=0.99)
results.loess_high_missense_uncon <- loess(forloess_missense_shet2bins_cadd6bins_high_uncon$obs_exp ~ forloess_missense_shet2bins_cadd6bins_high_uncon$midpoint, weights = forloess_missense_shet2bins_cadd6bins_high_uncon$n, span=0.99)
results.loess_low_missense_uncon <- loess(forloess_missense_shet2bins_cadd6bins_low_uncon$obs_exp ~ forloess_missense_shet2bins_cadd6bins_low_uncon$midpoint, weights = forloess_missense_shet2bins_cadd6bins_low_uncon$n, span=0.99)

results_loess_combined_missense <- data.frame(x = c(results.loess_high_missense_con$x, results.loess_low_missense_con$x, results.loess_high_missense_uncon$x, results.loess_low_missense_uncon$x),
           fitted = c(results.loess_high_missense_con$fitted, results.loess_low_missense_con$fitted, results.loess_high_missense_uncon$fitted, results.loess_low_missense_uncon$fitted),
           shetbin = c(rep('high', 6), rep('low', 6), rep('high', 6), rep('low', 6)),
           constrained = c(rep('True', 12), rep('False', 12)))

ggplot(rbind(forloess_missense_shet2bins_cadd6bins_high_con, forloess_missense_shet2bins_cadd6bins_low_con, forloess_missense_shet2bins_cadd6bins_high_uncon, forloess_missense_shet2bins_cadd6bins_low_uncon), aes(x = midpoint, y = obs_exp, shape=constrained, color=shetbin, lty=constrained)) +
  geom_point() +
  geom_pointrange(aes(ymin=ci.lower, ymax=ci.upper)) +
  theme_classic() +
  geom_hline(yintercept = 1, color='gray', lty=2) +
  labs(x = "CADD", y = 'Enrichment (obs/exp missense)') +
  scale_color_manual(values = c(mypal[2], mypal[1])) +
  geom_line(aes(y=results_loess_combined_missense$fitted, x=results_loess_combined_missense$x))
```

### Nonsense weights  

We split nonsense variants by sHet (high vs low), and by four bins of CADD scores for high sHet genes and seven bins for low sHet genes. As before, we found similar results when varying the number of CADD bins. This is the section that will need to be modified heavily for different datasets (e.g. fewer bins if the datasets are smaller).  

```{r function for 7 bins of nonsense}
get_expected_by_shet2bins_cadd7bins_nonsense <- function(rate_group) {
  return_rates <- rate_group %>%
    filter(cq=='nonsense') %>%
    mutate(sex_chr = ifelse(chrom=='X', 'sex', 'auto')) %>%
    mutate(shetbin = ifelse(shethigh=='TRUE', 'high', 'low')) %>%
    mutate(cadd_bin = ifelse(shetbin=='high',
                           case_when(score < 15 ~ 1,
                              score < 30 & score >= 15 ~ 3,
                              score < 45 & score >= 30 ~ 5,
                              score >= 45 ~ 7),
                           case_when(score < 7.5 ~ 1,
                              score < 15 & score >= 7.5 ~ 2,
                              score < 22.5 & score >= 15 ~ 3,
                              score < 30 & score >= 22.5 ~ 4,
                              score < 37.5 & score >= 30 ~ 5,
                              score < 45 & score >= 37.5 ~ 6,
                              score >= 45 ~ 7))) %>%
    filter(!is.na(cadd_bin)) %>%
    group_by(cadd_bin, shetbin, sex_chr) %>%
    summarize(sum_prob = sum(prob)) %>%
    mutate(expected = ifelse(sex_chr=='auto', sum_prob*autosomal, sum_prob*autosomal*x_factor)) %>%
    group_by(cadd_bin, shetbin) %>%
    summarize(total_expected = sum(expected))
  
  return(return_rates)
}
```

```{r nonsense}
## observed
obs_nonsense <- all_wweights %>%
    filter(cq=='stop_gained') %>%
    mutate(shetbin = ifelse(shethigh=='TRUE', 'high', 'low')) %>%
    mutate(cadd_bin = ifelse(shetbin=='high',
                           case_when(score < 15 ~ 1,
                              score < 30 & score >= 15 ~ 3,
                              score < 45 & score >= 30 ~ 5,
                              score >= 45 ~ 7),
                           case_when(score < 7.5 ~ 1,
                              score < 15 & score >= 7.5 ~ 2,
                              score < 22.5 & score >= 15 ~ 3,
                              score < 30 & score >= 22.5 ~ 4,
                              score < 37.5 & score >= 30 ~ 5,
                              score < 45 & score >= 37.5 ~ 6,
                              score >= 45 ~ 7))) %>%
    filter(!is.na(cadd_bin)) %>%
    group_by(cadd_bin, shetbin) %>%
    tally() %>%
    mutate(midpoint = case_when(cadd_bin==1 ~ 3.75,
                              cadd_bin==2 ~ 3.75+7.5,
                              cadd_bin==3 ~ 3.75+15,
                              cadd_bin==4 ~ 3.75+22.5,
                              cadd_bin==5 ~ 3.75+30,
                              cadd_bin==6 ~ 3.75+37.5,
                              cadd_bin==7 ~ 3.75+45))

## expected
exp_nonsense_rates1 <- get_expected_by_shet2bins_cadd7bins_nonsense(rates_1)
exp_nonsense_rates2 <- get_expected_by_shet2bins_cadd7bins_nonsense(rates_2)
exp_nonsense_rates3 <- get_expected_by_shet2bins_cadd7bins_nonsense(rates_3)
exp_nonsense_rates4 <- get_expected_by_shet2bins_cadd7bins_nonsense(rates_4)
exp_nonsense_rates5 <- get_expected_by_shet2bins_cadd7bins_nonsense(rates_5)

combined_exp_nonsense_rates <- bind_rows(exp_nonsense_rates1,
                                         exp_nonsense_rates2,
                                         exp_nonsense_rates3,
                                         exp_nonsense_rates4,
                                         exp_nonsense_rates5) %>%
    replace(., is.na(.), 0) %>%
    group_by(cadd_bin, shetbin) %>%
    summarize_all(sum)
  
## combine observed and expected
obs_exp_nonsense <- obs_nonsense %>%
    left_join(combined_exp_nonsense_rates, by=c('cadd_bin', 'shetbin')) %>%
    mutate(obs_exp = n/total_expected)
  
## add confidence intervals
obs_exp_nonsense$ci.lower <- sapply(obs_exp_nonsense$n, function(x) poisson.test(x, conf.level = 0.95)$conf.int[1])/obs_exp_nonsense$total_expected
obs_exp_nonsense$ci.upper <- sapply(obs_exp_nonsense$n, function(x) poisson.test(x, conf.level = 0.95)$conf.int[2])/obs_exp_nonsense$total_expected

ggplot(obs_exp_nonsense, aes(x = midpoint, y = obs_exp, group=shetbin, color=shetbin)) +
  geom_point() +
  geom_line() +
  geom_pointrange(aes(ymin=ci.lower, ymax=ci.upper)) +
  theme_classic() +
  geom_hline(yintercept = 1, color='gray', lty=2) +
  labs(x = "CADD", y = 'Enrichment (obs/exp nonsense)') +
  scale_color_manual(values = c(mypal[2], mypal[1]))
```

```{r run loess for nonsense variants}
## loess on high vs low separately
forloess_nonsense_high <- obs_exp_nonsense %>% filter(shetbin=='high')
forloess_nonsense_low <- obs_exp_nonsense %>% filter(shetbin=='low')

results.loess_high_nonsense <- loess(forloess_nonsense_high$obs_exp ~ forloess_nonsense_high$midpoint, weights = forloess_nonsense_high$n, span=1)
results.loess_low_nonsense <- loess(forloess_nonsense_low$obs_exp ~ forloess_nonsense_low$midpoint, weights = forloess_nonsense_low$n, span=1)
# only the first produced NaNs
  
results_loess_combined_nonsense <- data.frame(x = c(results.loess_high_nonsense$x,
                                                    results.loess_low_nonsense$x),
                                              fitted = c(results.loess_high_nonsense$fitted,
                                                        results.loess_low_nonsense$fitted),
                                              shetbin = c(rep('high', 4), rep('low', 7)))

## plotting loess
ggplot(rbind(forloess_nonsense_high, forloess_nonsense_low), aes(x = midpoint, y = obs_exp, group=shetbin, color=shetbin)) +
  geom_point() +
  geom_pointrange(aes(ymin=ci.lower, ymax=ci.upper)) +
  theme_classic() +
  geom_hline(yintercept = 1, color='gray', lty=2) +
  labs(x = "CADD", y = 'Enrichment (obs/exp nonsense)') +
  scale_color_manual(values = c(mypal[2], mypal[1])) +
  geom_line(aes(y=results_loess_combined_nonsense$fitted, x=results_loess_combined_nonsense$x))
```

## Run loess and make weight dictionaries   

```{r function to make loess dictionary}
get_loess_dic <- function(loess_high_con, loess_high_uncon, loess_low_con, loess_low_uncon, cq){
  vals <- seq(0,52.5,0.001)
  n_rows <- nrow(vals)
  if(cq=='nonsense'){
    # all "con" will be set to NA
    wdic_shethigh <- data.frame(cq = cq,
                                score = vals,
                                shetbin = 'high',
                                con = 'NA',
                                uncon = predict(loess_high_uncon, vals))
    lowest_uncon_value_shethigh <- min(wdic_shethigh$uncon, na.rm=T)
    lowest_score_lowest_uncon_value_shethigh <- min(wdic_shethigh$score[wdic_shethigh$uncon <= lowest_uncon_value_shethigh & !is.na(wdic_shethigh$uncon)])
    wdic_shethigh$uncon[wdic_shethigh$score < lowest_score_lowest_uncon_value_shethigh] <- lowest_uncon_value_shethigh
    highest_uncon_value_shethigh <- max(wdic_shethigh$uncon, na.rm=T)
    highest_score_highest_uncon_value_shethigh <- max(wdic_shethigh$score[wdic_shethigh$uncon >= highest_uncon_value_shethigh & !is.na(wdic_shethigh$uncon)])
    wdic_shethigh$uncon[wdic_shethigh$score > highest_score_highest_uncon_value_shethigh] <- highest_uncon_value_shethigh
    
    wdic_shetlow <- data.frame(cq = cq,
                               score = vals,
                               shetbin = 'low',
                               con = 'NA',
                               uncon = predict(loess_low_uncon, vals))
    lowest_uncon_value_shetlow <- min(wdic_shetlow$uncon, na.rm=T)
    lowest_score_lowest_uncon_value_shetlow <- min(wdic_shetlow$score[wdic_shetlow$uncon <= lowest_uncon_value_shetlow & !is.na(wdic_shetlow$uncon)])
    wdic_shetlow$uncon[wdic_shetlow$score < lowest_score_lowest_uncon_value_shetlow] <- lowest_uncon_value_shetlow
    highest_uncon_value_shetlow <- max(wdic_shetlow$uncon, na.rm=T)
    highest_score_highest_uncon_value_shetlow <- max(wdic_shetlow$score[wdic_shetlow$uncon >= highest_uncon_value_shetlow & !is.na(wdic_shetlow$uncon)])
    wdic_shetlow$uncon[wdic_shetlow$score > highest_score_highest_uncon_value_shetlow] <- highest_uncon_value_shetlow

    wdic_full <- rbind(wdic_shethigh, wdic_shetlow)
  } else if(cq=='missense'){
    # all four needed
    wdic_shethigh <- data.frame(cq = cq,
                                score = vals,
                                shetbin = 'high',
                                con = predict(loess_high_con, vals),
                                uncon = predict(loess_high_uncon, vals))
    lowest_uncon_value_shethigh <- min(wdic_shethigh$uncon, na.rm=T)
    lowest_score_lowest_uncon_value_shethigh <- min(wdic_shethigh$score[wdic_shethigh$uncon <= lowest_uncon_value_shethigh & !is.na(wdic_shethigh$uncon)])
    wdic_shethigh$uncon[wdic_shethigh$score < lowest_score_lowest_uncon_value_shethigh] <- lowest_uncon_value_shethigh
    highest_uncon_value_shethigh <- max(wdic_shethigh$uncon, na.rm=T)
    highest_score_highest_uncon_value_shethigh <- max(wdic_shethigh$score[wdic_shethigh$uncon >= highest_uncon_value_shethigh & !is.na(wdic_shethigh$uncon)])
    wdic_shethigh$uncon[wdic_shethigh$score > highest_score_highest_uncon_value_shethigh] <- highest_uncon_value_shethigh

    lowest_con_value_shethigh <- min(wdic_shethigh$con, na.rm=T)
    lowest_score_lowest_con_value_shethigh <- min(wdic_shethigh$score[wdic_shethigh$con <= lowest_con_value_shethigh & !is.na(wdic_shethigh$con)])
    wdic_shethigh$con[wdic_shethigh$score < lowest_score_lowest_con_value_shethigh] <- lowest_con_value_shethigh
    highest_con_value_shethigh <- max(wdic_shethigh$con, na.rm=T)
    highest_score_highest_con_value_shethigh <- max(wdic_shethigh$score[wdic_shethigh$con >= highest_con_value_shethigh & !is.na(wdic_shethigh$con)])
    wdic_shethigh$con[wdic_shethigh$score > highest_score_highest_con_value_shethigh] <- highest_con_value_shethigh

    wdic_shetlow <- data.frame(cq = cq,
                               score = vals,
                               shetbin = 'low',
                               con = predict(loess_low_con, vals),
                               uncon = predict(loess_low_uncon, vals))
    # define lowest uncon value
    lowest_uncon_value_shetlow <- min(wdic_shetlow$uncon, na.rm=T)
    # define lowest score associated with that value
    lowest_score_lowest_uncon_value_shetlow <- min(wdic_shetlow$score[wdic_shetlow$uncon <= lowest_uncon_value_shetlow & !is.na(wdic_shetlow$uncon)])
    # for anything with a score lower than this, make sure they are set to the lowest value
    wdic_shetlow$uncon[wdic_shetlow$score < lowest_score_lowest_uncon_value_shetlow] <- lowest_uncon_value_shetlow
    
    # define the highest uncon value
    highest_uncon_value_shetlow <- max(wdic_shetlow$uncon, na.rm=T)
    # define highest score associated with that value
    highest_score_highest_uncon_value_shetlow <- max(wdic_shetlow$score[wdic_shetlow$uncon >= highest_uncon_value_shetlow & !is.na(wdic_shetlow$uncon)])
    # for anything with a score higher than this, make sure they are set to the highest value
    wdic_shetlow$uncon[wdic_shetlow$score > highest_score_highest_uncon_value_shetlow] <- highest_uncon_value_shetlow

    # repeat for con
    lowest_con_value_shetlow <- min(wdic_shetlow$con, na.rm=T)
    lowest_score_lowest_con_value_shetlow <- min(wdic_shetlow$score[wdic_shetlow$con <= lowest_con_value_shetlow & !is.na(wdic_shetlow$con)])
    wdic_shetlow$con[wdic_shetlow$score < lowest_score_lowest_con_value_shetlow] <- lowest_con_value_shetlow
    highest_con_value_shetlow <- max(wdic_shetlow$con, na.rm=T)
    highest_score_highest_con_value_shetlow <- max(wdic_shetlow$score[wdic_shetlow$con >= highest_con_value_shetlow & !is.na(wdic_shetlow$con)])
    wdic_shetlow$con[wdic_shetlow$score > highest_score_highest_con_value_shetlow] <- highest_con_value_shetlow
    
    wdic_full <- rbind(wdic_shethigh, wdic_shetlow)
    # set all rates below 1 to 1 for con
    wdic_full$con[wdic_full$con < 1] <- 1
  }
  # set all rates below 1 to 1
  wdic_full$uncon[wdic_full$uncon < 1] <- 1
  
  return(wdic_full)
}
```

```{r make OR weight dictionaries}
weightdic_nonsense <- get_loess_dic(loess_high_con = NA,
                                    loess_high_uncon = results.loess_high_nonsense,
                                    loess_low_con = NA,
                                    loess_low_uncon = results.loess_low_nonsense,
                                    cq = "nonsense")

weightdic_missense <- get_loess_dic(loess_high_con = results.loess_high_missense_con,
                                    loess_high_uncon = results.loess_high_missense_uncon,
                                    loess_low_con = results.loess_low_missense_con,
                                    loess_low_uncon = results.loess_low_missense_uncon,
                                    cq = "missense")
```

```{r frameshift}
## frameshift based on the nonsense values
frameshift <- data.frame(cq = "frameshift",
                score = NA,
                shetbin = c('high', 'low'),
                con = NA,
                conlower = NA,
                conupper = NA,
                uncon = weightdic_nonsense %>% group_by(shetbin) %>% summarize(maxweight = max(uncon)) %>% pull(maxweight),
                unconlower = NA,
                unconupper = NA)
```

```{r splice lof}
## for splice_lof variants, need to determine enrichments by shet categories
obs_slof_shet2bins <- all_wweights %>%
  filter(cq %in% c('splice_acceptor_variant', 'splice_donor_variant') & size==0) %>%
  mutate(shetbin = ifelse(shethigh=='TRUE', 'high', 'low')) %>%
  group_by(shetbin) %>%
  tally()

shet2bins_slof_exp_rates1 <- get_expected_by_shet2bins_bycq(rates_1, 'splice_lof')
shet2bins_slof_exp_rates2 <- get_expected_by_shet2bins_bycq(rates_2, 'splice_lof')
shet2bins_slof_exp_rates3 <- get_expected_by_shet2bins_bycq(rates_3, 'splice_lof')
shet2bins_slof_exp_rates4 <- get_expected_by_shet2bins_bycq(rates_4, 'splice_lof')
shet2bins_slof_exp_rates5 <- get_expected_by_shet2bins_bycq(rates_5, 'splice_lof')

combined_shet2bins_slof_exp_rates <- bind_rows(shet2bins_slof_exp_rates1,
                                                shet2bins_slof_exp_rates2,
                                                shet2bins_slof_exp_rates3,
                                                shet2bins_slof_exp_rates4,
                                                shet2bins_slof_exp_rates5) %>%
  replace(., is.na(.), 0) %>%
  group_by(shetbin) %>%
  summarize_all(sum)

obs_exp_shet2bins_slof <- obs_slof_shet2bins %>%
  left_join(combined_shet2bins_slof_exp_rates, by=c('shetbin')) %>%
  mutate(obs_exp = n/total_expected)

# add confidence intervals
obs_exp_shet2bins_slof$ci.lower <- sapply(obs_exp_shet2bins_slof$n, function(x) poisson.test(x, conf.level = 0.95)$conf.int[1])/obs_exp_shet2bins_slof$total_expected
obs_exp_shet2bins_slof$ci.upper <- sapply(obs_exp_shet2bins_slof$n, function(x) poisson.test(x, conf.level = 0.95)$conf.int[2])/obs_exp_shet2bins_slof$total_expected

splice_lof <- data.frame(cq = "splice_lof",
           score = NA,
           shetbin = c('high', 'low'),
           con = NA,
           conlower = NA,
           conupper = NA,
           uncon = obs_exp_shet2bins_slof %>% pull(obs_exp),
           unconlower = obs_exp_shet2bins_slof %>% pull(ci.lower),
           unconupper = obs_exp_shet2bins_slof %>% pull(ci.upper))
```

```{r inframe}
## for inframe variants, need to determine enrichments by shet categories
obs_mis_shet2bins <- all_wweights %>%
  filter(cq=='missense_variant' & size==0) %>%
  mutate(shetbin = ifelse(shethigh=='TRUE', 'high', 'low')) %>%
  group_by(shetbin) %>%
  tally()

shet2bins_mis_exp_rates1 <- get_expected_by_shet2bins_bycq(rates_1, 'missense')
shet2bins_mis_exp_rates2 <- get_expected_by_shet2bins_bycq(rates_2, 'missense')
shet2bins_mis_exp_rates3 <- get_expected_by_shet2bins_bycq(rates_3, 'missense')
shet2bins_mis_exp_rates4 <- get_expected_by_shet2bins_bycq(rates_4, 'missense')
shet2bins_mis_exp_rates5 <- get_expected_by_shet2bins_bycq(rates_5, 'missense')

combined_shet2bins_mis_exp_rates <- bind_rows(shet2bins_mis_exp_rates1,
                                                shet2bins_mis_exp_rates2,
                                                shet2bins_mis_exp_rates3,
                                                shet2bins_mis_exp_rates4,
                                                shet2bins_mis_exp_rates5) %>%
  replace(., is.na(.), 0) %>%
  group_by(shetbin) %>%
  summarize_all(sum)

obs_exp_shet2bins_mis <- obs_mis_shet2bins %>%
  left_join(combined_shet2bins_mis_exp_rates, by=c('shetbin')) %>%
  mutate(obs_exp = n/total_expected)

# add confidence intervals
obs_exp_shet2bins_mis$ci.lower <- sapply(obs_exp_shet2bins_mis$n, function(x) poisson.test(x, conf.level = 0.95)$conf.int[1])/obs_exp_shet2bins_mis$total_expected
obs_exp_shet2bins_mis$ci.upper <- sapply(obs_exp_shet2bins_mis$n, function(x) poisson.test(x, conf.level = 0.95)$conf.int[2])/obs_exp_shet2bins_mis$total_expected

inframe <- data.frame(cq = 'inframe',
                      score = NA,
                      shetbin = c('high', 'low'),
                      con = NA,
                      conlower = NA,
                      conupper = NA,
                      uncon = obs_exp_shet2bins_mis %>% pull(obs_exp),
                      unconlower = NA,
                      unconupper = NA)
```

```{r combine and plot other weights}
weightdic_other <- rbind(synon, inframe, splice_lof, frameshift)

ggplot(weightdic_other, aes(x = cq, y = uncon, color=shetbin)) +
  geom_point(size = 3) +
  geom_pointrange(aes(ymin=unconlower, ymax=unconupper)) +
  theme_classic() +
  geom_hline(yintercept = 1, color='gray', lty=2) +
  labs(x = "CADD", y = 'Enrichment (obs/exp)') +
  scale_color_manual(values = c(mypal[2], mypal[1]))
```

## Positive predictive values (PPV)  

```{r ppv}
# combine weight dics
weightdic_combined <- rbind(weightdic_other[,c("cq","score", "shetbin", "con","uncon")], weightdic_missense, weightdic_nonsense)

#get ppv not odds ratio
ppv <- weightdic_combined

# previously defined as minimum of con and uncon (pre-shet split)
sub <- ppv %>% summarize(sub = min(con, uncon, na.rm=T)) %>% pull(sub)
sub <- as.numeric(sub)
conOR <- as.numeric(weightdic_combined$con) -sub + 1 
unconOR <- as.numeric(weightdic_combined$uncon) -sub + 1 

ppv$con <- (conOR-1)/conOR
ppv$uncon <- (unconOR-1)/unconOR

ppv$uncon[ppv$cq=="synonymous"]<- 0.001
```

If you want to recreate part of our supplemental plot on PPV weights, you can run the following chunk of code.  

```{r make ppv plots}
library(tidyr)
library(cowplot)

other_ppv.plot <- ggplot(ppv %>% filter(!(cq %in% c('missense', 'nonsense'))), aes(x = cq, y = uncon, color=shetbin)) +
  geom_point(size = 3) +
  scale_color_manual(values = c(mypal[2], mypal[1]), labels = c('high shet', 'low shet')) +
  theme_classic() +
  labs(x = "", y = 'PPV', title = '') +
  ylim(0,1) +
  theme(legend.position = 'none', plot.title = element_text(hjust = 0.5))

missense_ppv.plot <- ggplot(ppv %>% filter(cq=='missense' & (score %% 1)==0) %>% gather(con, uncon, key = 'constrained', value = 'ppv'),
       aes(x = score, y = ppv, color=shetbin, lty = constrained)) +
  geom_line() +
  theme_classic() +
  labs(x = "CADD", y = 'PPV', title = 'missense', lty = 'In MCR', color= "") +
  scale_color_manual(values = c(mypal[2], mypal[1]), labels = c('high shet', 'low shet')) +
  scale_linetype_manual(values = c(2,1), labels = c('True', 'False')) +
  ylim(0,1) +
  xlim(c(0,35)) +
  theme(legend.position = 'none', plot.title = element_text(hjust = 0.5))

nonsense_ppv.plot <- ggplot(ppv %>% filter(cq=='nonsense' & score <= 48.75), aes(x = score, y = uncon, color=shetbin)) +
  geom_line() +
  theme_classic() +
  labs(x = "CADD", y = 'PPV', title = 'nonsense') +
  scale_color_manual(values = c(mypal[2], mypal[1]), labels = c('high shet', 'low shet')) +
  ylim(0,1) +
  xlim(c(0,50)) +
  theme(legend.position = 'none', plot.title = element_text(hjust = 0.5))

plot_grid(other_ppv.plot, missense_ppv.plot, nonsense_ppv.plot, nrow=1)
```

To save a file in the same order as the one we released, run the following code. You may want to rename the file.  

```{r save files, eval=FALSE}
reorder_ppv <- ppv %>%
  mutate(shetbin = ifelse(shetbin=='high', 'shethigh', 'shetlow')) %>%
  gather(con, uncon, key = 'constrained', value = 'testcol') %>%
  spread(shetbin, testcol) %>%
  mutate(constrained = ifelse(constrained=='con', 'True', 'False')) %>%
  filter(is.na(score) | score <= 50)

write.table(reorder_ppv, file = "weights_ppv.tab", sep = "\t", row.names = F, quote = F)
```


